Effects of Reynolds Number on the Energy Conversion and Near-Wake Dynamics of a High Solidity Vertical-Axis Cross-Flow Turbine

### 2015-10-12 paper.tex
Better summary for Re effects on static foil perf

Hydro- or airfoil section performance is generally characterized by the
lift-to-drag force ratio $L/D$  (or lift vs drag coefficient, $c_L/c_D$ ), which
is primarily a function of the foil's angle of attack $\alpha$. The angles of
attack for cross-flow turbine blades are a function of turbine tip speed ratio,
induction (slowing/turning of the free stream flow prior to reaching the
turbine), and turbine rotation or azimuthal angle $\theta$. Looking at static
airfoil data, we see that the static stall angle---the angle just beyond that at
which $L/D$ is maximum---increases with blade chord Reynolds number $Re_c$
(reference?). This is caused by the boundary layer on the foil becoming more
turbulent and more energetic, which enables it to advance further downstream
against the adverse pressure gradient on the suction side and delay separation
to higher angles of attack, as well as diminish the negative effects of a
laminar separation bubble occurring near the leading edge (reference?). For
smooth airfoils, this transition can cause a dramatic increase in foil
performance, at a blade chord Reynolds number on the order of $10^5$
\cite{McMasters1980}, shown in Figure~\ref{fig:McMasters}. A review of low
Reynolds number airfoil behavior is presented in \cite{Lissaman1983}. Note that
there is a distinct lack of highly reliable data for airfoils in this
transitional regime and below. An evaluation of the various databases relevant
to cross-flow turbines is presented in \cite{Bedon2014}.

\todo[inline]{Laminar separation bubble - is it always there? also need
reference}

>>>>

Hydro- or airfoil section performance is generally characterized by the
lift-to-drag force ratio $L/D$  (or lift vs drag coefficient, $c_L/c_D$ ), which
is primarily a function of the foil's angle of attack $\alpha$. The angles of
attack for cross-flow turbine blades are a function of turbine tip speed ratio,
induction (slowing/turning of the free stream flow prior to reaching the
turbine), and turbine rotation or azimuthal angle $\theta$. Looking at static
airfoil data, we see that the static stall angle---the angle just beyond that at
which $L/D$ is maximum---increases with blade chord Reynolds number $Re_c$
\cite{Jacobs1937}.

A review of Reynolds number effects on airfoil behavior is presented in
\cite{Lissaman1983}. In general, airfoil performance is enhanced as the boundary
layer on the foil transitions to turbulence closer to the leading edge, which
enables it to advance further downstream against the adverse pressure gradient
on the suction side, delaying separation to higher angles of attack. For smooth
airfoils, this transition can cause a dramatic increase in foil performance, at
a blade chord Reynolds number on the order of $10^5$ \cite{McMasters1980}, shown
in Figure~\ref{fig:McMasters}. Note that there is a distinct lack of highly
reliable data for airfoils in this transitional regime and below. An evaluation
of the various databases relevant to cross-flow turbines is presented in
\cite{Bedon2014}.
